---
title: "Sample Summary Report - Capital Budgeting"
author: "K P Joseph"
date: '`r Sys.Date()`'
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(shiny)
library(htmltools)
library(purrr)
library(leaflet)
library(highcharter)
library(zoo)
library(DT)
library(sf)
library(tigris)
library(stringr)
library(xml2)
library(shinyWidgets)
library(FinancialMath)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE,message = FALSE)
```

## Project Selection - Introduction

Firms do not often have unlimited resources to fund all the projects in their portfolio. In Capital Budgeting, firms can use the NPV rule to select the projects to include in their current budget. Any project with negative NPV is removed from the budget. 

This is a sample project selection process created in R.

Please download the [Rmarkdown file from this link](https://raw.githubusercontent.com/peekpetAnalystJourney/peekpetAnalystJourney/main/app.Rmd)

#### Key Assumptions
1. Production profile assumes 10% annualized decline	

```{r }
url_prod   <- "https://raw.githubusercontent.com/peekpetAnalystJourney/peekpetAnalystJourney/main/Production.csv"

raw_prod   <- readr::read_csv(url_prod)
raw_prod %>% 
  tidyr::pivot_longer(-Year, names_to = "series",values_to = "value") %>% 
  ggplot2::ggplot(aes(x=Year,y=value,fill=series))+geom_area()+
  labs(title = "Annual Oil Production Volume bbl",y="oil volume bbl")

raw_prod %>% 
  tidyr::pivot_longer(-Year, names_to = "series",values_to = "value") %>% 
 dplyr::group_by(series) %>% 
 dplyr::summarise("Reserve MMbbl"=sum(value)/1e6) %>% 
  knitr::kable(caption = "Project Reserve",digits=2)
```

2. WCS price is derived from WTI future price - dynamically linked via Web scrapping from -  ([source](https://www.oilcrudeprice.com/wti-oil-price-future/)) and a constant WCS differential of USD18/bbl. Price set as fixed value till from 2032 onward.

Canadian Crudes are mostly sold at Western Canadian Select (WCS) crude oil price.

0.5 and 1.5 multiplier used to generate low and high price forecast respectively.

```{r comment=FALSE,warning=FALSE}
wcs_diff <- 18
url_price <- "https://www.oilcrudeprice.com/wti-oil-price-future/"

price <- xml2::read_html(url_price)
price <- price %>% rvest::html_nodes(css="table.tables") %>% rvest::html_table()


# binding columns together
df <- do.call(rbind, price)

price <- as.data.frame(df) %>% 
  dplyr::select(Contract,Last) %>% 
  tidyr::separate(Contract,into=c("month","Year")) %>% 
    dplyr::select(Year,Last) %>% 
    dplyr::rename(Base_Price1 = Last) %>%
  dplyr::mutate(Base_Price2 = Base_Price1 - wcs_diff)%>% 
  dplyr::mutate(Low_Price2 = Base_Price2*0.5,High_Price2 = Base_Price2*1.5) %>% 
  dplyr::filter(Year>2022&Year<2032) %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarise(Base_Price=mean(Base_Price2),Low_Price=mean(Low_Price2),High_Price=mean(High_Price2)) 

fixed_base = price$Base_Price[9]
fixed_high = 39*1.5
fixed_low = 39*0.5

temp <- tibble("Year"=seq(2032,2039,1),"Base_Price"=fixed_base,"Low_Price"=fixed_low,"High_Price"=fixed_high)

  price <- rbind(price,temp)


price %>% 
  tidyr::pivot_longer(-Year,names_to = 'series',values_to = 'value') %>% 
  ggplot(aes(Year,value,color=series))+geom_point() +
  labs(title = "Range of Oil Price Assumed",y="WCS Oil Price US$/bbl")
```


						
3. Average Opex per barrel for Canadian operator from [Source](https://knoema.com/rqaebad/cost-of-producing-a-barrel-of-crude-oil-by-country)


4. Fx dynamically linked via Web scrapping from [xe.com](https://www.xe.com/currencyconverter/convert/?Amount=1&From=USD&To=CAD)

```{r}
fx_url = "https://www.xe.com/currencyconverter/convert/?Amount=1&From=USD&To=CAD"
css_pth = "#__next > div:nth-child(2) > div.fluid-container__BaseFluidContainer-qoidzu-0.gJBOzk > section > div:nth-child(2) > div > main > form > div:nth-child(2) > div.result__Repulsor-sc-1bsijpp-4.dZGBTm > div:nth-child(1) > div.unit-rates___StyledDiv-sc-1dk593y-0.dEqdnx > p"
fx <- rvest::read_html(fx_url)
fx <- fx %>% rvest::html_element(css = css_pth) %>% rvest::html_text()
fx <- str_split(fx," ",n=5,simplify = TRUE)[4]
fx <- as.numeric(fx)
```


```{r}
#Key assumption
Tax_rate = 0.2
disc_factor = 0.1
opex_per_barrel = 23.3

df1 <- tibble(Input=c("fx: CA $/US $",	
"Average Opex per barrel (USD/bbl)",	
"Tax Rate",	
"Discount factor"),values =c(fx,opex_per_barrel,sprintf("%.0f%%",Tax_rate*100),sprintf("%.0f%%",disc_factor*100))) 
df1 %>% 
  knitr::kable(caption = "Basic Input Parameters",digits=3)
```


6. These projects are incremental projects for the mining and upgrading facility. Mining will continue after these projects ended. No mine financial deposit security required during the project life.			




```{r}
df2 <- tibble("Project"=names(raw_prod[2:5]),Cost=c(10000000,
9640265,
12180473,
15187827)) 

df2 %>% 
  ggplot(aes(x=Project,y=Cost,fill=Project))+geom_col()+labs(title = "Assumed Upfront Project Cost",y="Project Cost, CA$")+
  annotate("text",x="Project_1",y=df2$Cost[1]/2,angle=90, label=paste0(sprintf("CA$%.2f",df2$Cost[1]/1e6)," million"),color='white',size=7) +
  annotate("text",x="Project_2",y=df2$Cost[2]/2,angle=90, label=paste0(sprintf("CA$%.2f",df2$Cost[2]/1e6)," million"),color='white',size=7) +
  annotate("text",x="Project_3",y=df2$Cost[3]/2,angle=90, label=paste0(sprintf("CA$%.2f",df2$Cost[3]/1e6)," million"),color='white',size=7) +
  annotate("text",x="Project_4",y=df2$Cost[4]/2,angle=90, label=paste0(sprintf("CA$%.2f",df2$Cost[4]/1e6)," million"),color='white',size=7)
```

### Results

The results of the economics analysis is presented below and ranked using NPV.



```{r base}

df3 <- merge(raw_prod,price) %>% 
  dplyr::mutate(Project1_Revenue=Project_1*Base_Price / fx,
                Project2_Revenue=Project_2*Base_Price / fx,
                Project3_Revenue=Project_3*Base_Price / fx,
                Project4_Revenue=Project_4*Base_Price / fx,
                Project1_Opex=Project_1*opex_per_barrel / fx,
                Project2_Opex=Project_2*opex_per_barrel / fx,
                Project3_Opex=Project_3*opex_per_barrel / fx,
                Project4_Opex=Project_4*opex_per_barrel / fx,
                Project1_Cashflow=(Project1_Revenue - Project1_Opex),
                Project2_Cashflow=Project2_Revenue - Project2_Opex,
                Project3_Cashflow=Project3_Revenue - Project3_Opex,
                Project4_Cashflow=Project4_Revenue - Project4_Opex,
                Project1_Cashflow_AT=Project1_Cashflow*(1-Tax_rate),
                Project2_Cashflow_AT=Project2_Cashflow*(1-Tax_rate),
                Project3_Cashflow_AT=Project3_Cashflow*(1-Tax_rate),
                Project4_Cashflow_AT=Project4_Cashflow*(1-Tax_rate))

df4 <- tibble(Category=c("NPV (CA$ million)","IRR","Productivity Index"), 
       "Project 1"=c(FinancialMath::NPV(cf0=df2$Cost[1],cf=df3$Project1_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/1e6,
                     FinancialMath::IRR(cf0=df2$Cost[1],cf=df3$Project1_Cashflow_AT,times=seq(1,length(df3$Year),1))[2],
                     1+FinancialMath::NPV(cf0=df2$Cost[1],cf=df3$Project1_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/df2$Cost[1]),
       
       "Project 2"=c(FinancialMath::NPV(cf0=df2$Cost[2],cf=df3$Project2_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/1e6,
                     FinancialMath::IRR(cf0=df2$Cost[2],cf=df3$Project2_Cashflow_AT,times=seq(1,length(df3$Year),1))[2],
                     1+FinancialMath::NPV(cf0=df2$Cost[2],cf=df3$Project2_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/df2$Cost[2]),
       
       "Project 3"=c(FinancialMath::NPV(cf0=df2$Cost[3],cf=df3$Project3_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/1e6,
                     FinancialMath::IRR(cf0=df2$Cost[3],cf=df3$Project3_Cashflow_AT,times=seq(1,length(df3$Year),1))[2],
                     1+FinancialMath::NPV(cf0=df2$Cost[3],cf=df3$Project3_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/df2$Cost[3]),
       
       "Project 4"=c(FinancialMath::NPV(cf0=df2$Cost[4],cf=df3$Project4_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/1e6,
                     FinancialMath::IRR(cf0=df2$Cost[4],cf=df3$Project4_Cashflow_AT,times=seq(1,length(df3$Year),1))[2],
                     1+FinancialMath::NPV(cf0=df2$Cost[4],cf=df3$Project4_Cashflow_AT,times=seq(1,length(df3$Year),1),i=disc_factor)/df2$Cost[4])
      
) 

df4 %>% 
  knitr::kable(caption = "Summary of Results",digits = 2,booktabs=TRUE,longtable = TRUE) %>% 
  kableExtra::column_spec(4, color = 'white', background = 'green') %>% 
  kableExtra::column_spec(5, color = 'white', background = 'red') %>% 
  kableExtra::kable_styling("striped",full_width = TRUE)
  
head(df3) %>% 
  knitr::kable()
```



```{r low, include=FALSE}
df3a <- merge(raw_prod,price) %>% 
  dplyr::mutate(Project1_Revenue=Project_1*Low_Price / fx,
                Project2_Revenue=Project_2*Low_Price / fx,
                Project3_Revenue=Project_3*Low_Price / fx,
                Project4_Revenue=Project_4*Low_Price / fx,
                Project1_Opex=Project_1*opex_per_barrel / fx,
                Project2_Opex=Project_2*opex_per_barrel / fx,
                Project3_Opex=Project_3*opex_per_barrel / fx,
                Project4_Opex=Project_4*opex_per_barrel / fx,
                Project1_Cashflow=(Project1_Revenue - Project1_Opex),
                Project2_Cashflow=Project2_Revenue - Project2_Opex,
                Project3_Cashflow=Project3_Revenue - Project3_Opex,
                Project4_Cashflow=Project4_Revenue - Project4_Opex,
                Project1_Cashflow_AT=Project1_Cashflow*(1-Tax_rate),
                Project2_Cashflow_AT=Project2_Cashflow*(1-Tax_rate),
                Project3_Cashflow_AT=Project3_Cashflow*(1-Tax_rate),
                Project4_Cashflow_AT=Project4_Cashflow*(1-Tax_rate))

```


```{r high, include=FALSE}
df3b <- merge(raw_prod,price) %>% 
  dplyr::mutate(Project1_Revenue=Project_1*High_Price / fx,
                Project2_Revenue=Project_2*High_Price / fx,
                Project3_Revenue=Project_3*High_Price / fx,
                Project4_Revenue=Project_4*High_Price / fx,
                Project1_Opex=Project_1*opex_per_barrel / fx,
                Project2_Opex=Project_2*opex_per_barrel / fx,
                Project3_Opex=Project_3*opex_per_barrel / fx,
                Project4_Opex=Project_4*opex_per_barrel / fx,
                Project1_Cashflow=(Project1_Revenue - Project1_Opex),
                Project2_Cashflow=Project2_Revenue - Project2_Opex,
                Project3_Cashflow=Project3_Revenue - Project3_Opex,
                Project4_Cashflow=Project4_Revenue - Project4_Opex,
                Project1_Cashflow_AT=Project1_Cashflow*(1-Tax_rate),
                Project2_Cashflow_AT=Project2_Cashflow*(1-Tax_rate),
                Project3_Cashflow_AT=Project3_Cashflow*(1-Tax_rate),
                Project4_Cashflow_AT=Project4_Cashflow*(1-Tax_rate))

```


**Project 3** has the highest NPV and is therefore selected first. 

**Project 4** has negative NPV and is therefore removed from the budget. 

##### After-Tax Cashflow Plot per Project

```{r reactive-inputs}

df5 <- df3 %>% 
dplyr:: select(Year,ends_with("Cashflow_AT"))


# single selection

  ui = fluidPage(
    varSelectInput("variable", "Variable:", df5[2:5]),
    plotOutput("data")
    )

  server = function(input, output) {
    output$data <- renderPlot({
      ggplot(df5, aes(factor(Year),!!input$variable,fill=factor(Year))) + layer(geom='col',stat = "identity", position = "identity",show.legend = FALSE) + labs(title="Project Cashflow After Tax, CA$")
    })
  }
shinyApp(ui = ui, server = server)
```



### Sensitivity Results

##### Economics results for Base Oil Price at different Discount Factors


```{r}
ui <- fluidPage(
sliderInput(inputId = "Discf",
label = "Base Oil Price Scenario: Choose a Discount Factor",
value = 0.1, min = 0, max = 0.5),
tableOutput("data")
)
server <- function(input, output) {
  output$data <- renderTable({
  tibble(Category=c("NPV (CA$ million)","Productivity Index"), 
       "Project 1"=c(FinancialMath::NPV(cf0=df2$Cost[1],cf=df3$Project1_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[1],cf=df3$Project1_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/df2$Cost[1]),
       
       "Project 2"=c(FinancialMath::NPV(cf0=df2$Cost[2],cf=df3$Project2_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[2],cf=df3$Project2_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/df2$Cost[2]),
       
       "Project 3"=c(FinancialMath::NPV(cf0=df2$Cost[3],cf=df3$Project3_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[3],cf=df3$Project3_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/df2$Cost[3]),
       
       "Project 4"=c(FinancialMath::NPV(cf0=df2$Cost[4],cf=df3$Project4_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[4],cf=df3$Project4_Cashflow_AT,times=seq(1,length(df3$Year),1),i=input$Discf)/df2$Cost[4]) 
) 
    })
  }
shinyApp(ui = ui, server = server)
```


##### Economics results for High Oil Price at different Discount Factors


```{r}
ui <- fluidPage(
sliderInput(inputId = "Discf",
label = "High Oil Price Scenario: Choose a Discount Factor",
value = 0.1, min = 0, max = 0.5),
tableOutput("data")
)
server <- function(input, output) {
  output$data <- renderTable({
  tibble(Category=c("NPV (CA$ million)","Productivity Index"), 
       "Project 1"=c(FinancialMath::NPV(cf0=df2$Cost[1],cf=df3b$Project1_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[1],cf=df3b$Project1_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/df2$Cost[1]),
       
       "Project 2"=c(FinancialMath::NPV(cf0=df2$Cost[2],cf=df3b$Project2_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[2],cf=df3b$Project2_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/df2$Cost[2]),
       
       "Project 3"=c(FinancialMath::NPV(cf0=df2$Cost[3],cf=df3b$Project3_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[3],cf=df3b$Project3_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/df2$Cost[3]),
       
       "Project 4"=c(FinancialMath::NPV(cf0=df2$Cost[4],cf=df3b$Project4_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[4],cf=df3b$Project4_Cashflow_AT,times=seq(1,length(df3b$Year),1),i=input$Discf)/df2$Cost[4]) 
) 
    })
  }
shinyApp(ui = ui, server = server)
```


##### Economics results for Low Oil Price at different Discount Factors

At what discount factor can you invest any of these projects if the expected oil price outlook is in the low oil price range?


```{r}
ui <- fluidPage(
sliderInput(inputId = "Discf",
label = "Low Oil Price Scenario: Choose a Discount Factor",
value = 0.1, min = 0, max = 0.5),
tableOutput("data")
)
server <- function(input, output) {
  output$data <- renderTable({
  tibble(Category=c("NPV (CA$ million)","Productivity Index"), 
       "Project 1"=c(FinancialMath::NPV(cf0=df2$Cost[1],cf=df3a$Project1_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[1],cf=df3a$Project1_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/df2$Cost[1]),
       
       "Project 2"=c(FinancialMath::NPV(cf0=df2$Cost[2],cf=df3a$Project2_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[2],cf=df3a$Project2_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/df2$Cost[2]),
       
       "Project 3"=c(FinancialMath::NPV(cf0=df2$Cost[3],cf=df3a$Project3_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[3],cf=df3a$Project3_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/df2$Cost[3]),
       
       "Project 4"=c(FinancialMath::NPV(cf0=df2$Cost[4],cf=df3a$Project4_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/1e6,
                     1+FinancialMath::NPV(cf0=df2$Cost[4],cf=df3a$Project4_Cashflow_AT,times=seq(1,length(df3a$Year),1),i=input$Discf)/df2$Cost[4]) 
) 
    })
  }
shinyApp(ui = ui, server = server)
```



*Finally, revenue of oil and gas firms is very sensitive to oil price. Therefore, the expected crude price over the project life can influence the firm's commitment of capital to incremental projects. *

*The uncertainties in the cost of financing - which is often reflected in the discount factor - and exchange rates must be considered.*